### 2012-09-04 sections/implementation/execution-workflow.tex
moved acronym section, fixed filter-struct reference
fix two spelling errors

In order to be able to make comparisons on the field offsets of such a term,
the comparator needs to know the type of the comparison and the length of the
field offset. This information is parsed by execution engine once the query is
read and is therefore not available at compile time.  In order to subvert the
need to define complex branching statements, a dedicated comparator is defined
for every possible field length and comparison operation. A Python script
generates C source code for these comparators at compile time conforming to
the structure shown in listing \ref{filter-struct}. This allows the term
definitions to make runtime calls using a function name derived from the
combination of operation type and field length.

@@

The execution engine, as defined by the flow query language must read all the
flow records of a supplied trace into memory before starting the processing
pipeline.  Since, the filter stage uses the supplied set of absolute rules to
make a decision on whether or not to keep a flow record; it has to pass
through the whole in-memory recordset \emph{again} to produce the filter
results. This technique involves multiple linear runs on the trace and
therefore slows down when the ratio of number of filtered records to the total
number of flow-records is high. We optimized this behavior in \ac{NFQL} by
merging the filter stage with in-memory read of the trace. This means, a
decision on whether or not to make room for a record in memory and eventually
hold a pointer for it in filter results is done upfront as soon as the record
is read from the trace. In addition, if a request to write the filter stage
results to a flow-tools file has been made, the writes are also made as soon
the filter stage decision is available, thereby allowing
reading-filtering-writing to happen in $O(n)$ time, where $n$ is the number of
records in the trace. The filtered records are saved in a common location from
where they are referenced by each branch. This helps keep the memory costs at
a minimum when multiple branches are involved.  We used the publically
available Trace 7 from the SimpleWeb \cite{simpleweb} to compare the
performance of the optimized \ac{NFQL} implementation against the one defined
by the flow query language as shown in Fig.  \ref{fig:fv1-fv2-filter}. The
filter stage implementation with these optimizations runs 10 times faster and
is more pronounced on higher ratios.

@@

The merger is used to relate groups from different branches according to a
merging criterean. The implementation is not trivial since the number of
branches that need to be spawned is read from the query and is not known until
runtime. As a result, an iterator that can provide all possible permutations
of $m-$tuple (where $m$ is the number of branches) group record $ID$s was
needed.  The result of the iterator can then be used to make a match. The
merger stage begins by initializing this iterator passing it the number of
branches, and information about each branch. Then, it loops over to get a new
$m-$tuple of group record $ID$s on each iteration until the iterator returns
\texttt{false}.

>>>>

In order to be able to make comparisons on the field offsets of such a term,
the comparator needs to know the type of the comparison and the length of the
field offset. This information is parsed by execution engine once the query is
read and is therefore not available at compile time.  In order to subvert the
need to define complex branching statements, a dedicated comparator is defined
for every possible field length and comparison operation. A Python script
generates C source code for these comparators at compile time conforming to
the structure shown in listing \ref{lst:filter-struct}. This allows the term
definitions to make runtime calls using a function name derived from the
combination of operation type and field length.

@@

The execution engine, as defined by the flow query language must read all the
flow records of a supplied trace into memory before starting the processing
pipeline.  Since, the filter stage uses the supplied set of absolute rules to
make a decision on whether or not to keep a flow record; it has to pass
through the whole in-memory recordset \emph{again} to produce the filter
results. This technique involves multiple linear runs on the trace and
therefore slows down when the ratio of number of filtered records to the total
number of flow-records is high. We optimized this behavior in \ac{NFQL} by
merging the filter stage with in-memory read of the trace. This means, a
decision on whether or not to make room for a record in memory and eventually
hold a pointer for it in filter results is done upfront as soon as the record
is read from the trace. In addition, if a request to write the filter stage
results to a flow-tools file has been made, the writes are also made as soon
the filter stage decision is available, thereby allowing
reading-filtering-writing to happen in $O(n)$ time, where $n$ is the number of
records in the trace. The filtered records are saved in a common location from
where they are referenced by each branch. This helps keep the memory costs at
a minimum when multiple branches are involved.  We used the publicly
available Trace 7 from the SimpleWeb \cite{simpleweb} to compare the
performance of the optimized \ac{NFQL} implementation against the one defined
by the flow query language as shown in Fig.  \ref{fig:fv1-fv2-filter}. The
filter stage implementation with these optimizations runs 10 times faster and
is more pronounced on higher ratios.

@@

The merger is used to relate groups from different branches according to a
merging criterion. The implementation is not trivial since the number of
branches that need to be spawned is read from the query and is not known until
runtime. As a result, an iterator that can provide all possible permutations
of $m-$tuple (where $m$ is the number of branches) group record $ID$s was
needed.  The result of the iterator can then be used to make a match. The
merger stage begins by initializing this iterator passing it the number of
branches, and information about each branch. Then, it loops over to get a new
$m-$tuple of group record $ID$s on each iteration until the iterator returns
\texttt{false}.
